---
title: "HW2"
author: "M.Martinez-Raga"
date: "9/14/2020"
output: html_document
---

A single die has a sample space of 6 outcomes that are all mutually exclusive. For the purpose of this lab, we are focusing on calculating and testing the probability of the die being fixed toward 6. Theoretically, there is a 1/6, or 16.667%, probability of the die being judged unfair and a 5/6, or 83.333%, probability of the die being judged fair. 

Null hypothesis (Ho): Our die is fair, not fixed to 6.
Alternative hypothesis (Ha): Our die is unfair, fixed to 6.

______

PP1

For our first possible protocol, our number of trials (n) is 1. The rule is as follows: If we get a 6 after a single roll, we are deeming the die unfair. Any other number will be deemed fair. The probability of our null hypothesis being accepted or rejected are 5/6 and 1/6, respectively.

The standard error for this trial will be 0.3727, where n = 1 and p = 1/6.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
dice.roll <- sample(1:6, size = 1, replace = TRUE)
dice.roll
ifelse(dice.roll == 6, "UNFAIR","FAIR")
```

For PP1, we got a 6 in one roll, deeming the dice unfair. We accept the reject the null hypothesis for PP1.

______


PP2

For our second possible protocol, our number of trials (n) is 30. We used a 99% confidence interval to determine the range of occurences of 6 where the die would still be considered fair. Anything past that range, will be deemed unfair. We chose 99% confidence because the sample size is so small that any additional unit of deviation can misledingly deem the die unfair, therefore we want a more lenient range. 

Below we determine expected occurences of 6 in a trial of 30 rolls.
```{r}
theoretical_prob = 1/6
n_pp2 = 30
occurences_pp2 = theoretical_prob * n_pp2
occurences_pp2
```


Now, we determine standard error for this trial.
```{r}
se_pp2 = sqrt(n_pp2*(theoretical_prob*(1-theoretical_prob)))
se_pp2
```

In order to determine the range of fair occurences of 6 with confidence level of 99%, we use a z-score of 2.576.
```{r}
z_99 = 2.576
distance_pp2 = se_pp2 * z_99
floor_pp2 = occurences_pp2 - distance_pp2
ceiling_pp2 = occurences_pp2 + distance_pp2
range_pp2 = c(floor_pp2, ceiling_pp2)
range_pp2
```

If a fair die is rolled 30 times, the likely amount of occurences for 6 should be between 0 and 10 times. We will reject Ho if 6 is drawn above 10 times. *should we use 95% to standardize with other trials? z = 1.96, range = 1-9.*

Below is the graphical binomial distribution of PP2.
```{r}
success <- 0:30

plot(success,dbinom(success,size=30,prob=.1667),
     type='h',
     main='Binomial Distribution (n = 30, p = 1/6)',
     ylab='Probability',
     xlab ='# Occurences',
     lwd=3)
```
Hypothesis test PP2: 
```{r}
dice.roll <- sample(1:6, size = 30, replace = TRUE)
test_pp2 <- table(dice.roll)
test_pp2
```

Given an outcome of 4 occurences of 6, we accept the null hypothesis for PP2. All other sides also had an outcome within the range of 0 and 10.


______


PP3

For our third possible protocol, our number of trials (n) is 100. We used a 98% confidence interval to determine the range of occurences of 6 where the die would still be considered fair. Anything past that range, will be deemed unfair. We chose 98% confidence because the sample size is bigger than PP2, however we expect it to be far from the theoretical. 

Below we determine expected occurences of 6 in a trial of 100 rolls.
```{r}
theoretical_prob = 1/6
n_pp3 = 100
occurences_pp3 = theoretical_prob * n_pp3
occurences_pp3
```
Now, we determine standard error for this trial.
```{r}
se_pp3 = sqrt(n_pp3*(theoretical_prob*(1-theoretical_prob)))
se_pp3
```

In order to determine the range of fair occurences of 6 with confidence level of 98%, we use a z-score of 2.326.
```{r}
z_98 = 2.326
distance_pp3 = se_pp3 * z_98
floor_pp3 = occurences_pp3 - distance_pp3
ceiling_pp3 = occurences_pp3 + distance_pp3
range_pp3 = c(floor_pp3, ceiling_pp3)
range_pp3
```

If a fair die is rolled 100 times, the likely amount of occurences for 6 should be between 8 and 25 times. We will reject Ho if 6 is drawn above 25 times. *should we use 95% to standardize with other trials?*

Below is the graphical binomial distribution of PP3.
```{r}
success <- 0:100

plot(success,dbinom(success,size=100,prob=.1667),
     type='h',
     main='Binomial Distribution (n = 100, p = 1/6)',
     ylab='Probability',
     xlab ='# Occurences',
     lwd=3)
```

Hypothesis test PP3: 
```{r}
dice.roll <- sample(1:6, size = 100, replace = TRUE)
test_pp3 <- table(dice.roll)
test_pp3
```

Given an outcome of 12 occurences of 6, we accept the null hypothesis for PP3. In fact, all other sides had an outcome within the range of 8 and 25.

______

EP

For our experimental protocol, we want to roll our dice enough times so that 6 has occurences much closer to the theoretical mean of 16.7%. We have chosen to prove this by dramatically increasing the sample size. The confidence levels were also decreased proportionally to prove that the values  after time will continue staying within range. Each EP will be run 3 times. Our hypothesis remain the same. The following is the summary of EPs.
(EP1) n: 1,000, confidence: 95% 
(EP2) n: 10,000, 90%
(EP3) n: 100,000x, 68%


EP1
For EP1, we will repeat the procedure of our possible protocols, while changing rolls to 1,000 and our confidence level to 95%.

Below we determine expected occurences of 6 in a trial of 1,000 rolls.
```{r}
n_ep1 = 1000
occurences_ep1 = theoretical_prob * n_ep1
occurences_ep1
```
Now, we determine standard error for this trial.
```{r}
se_ep1 = sqrt(n_ep1*(theoretical_prob*(1-theoretical_prob)))
se_ep1
```

In order to determine the range of fair occurences of 6 with confidence level of 95%, we use a z-score of 1.96.
```{r}
z_95 = 1.96
distance_ep1 = se_ep1 * z_95
floor_ep1 = occurences_ep1 - distance_ep1
ceiling_ep1 = occurences_ep1 + distance_ep1
range_ep3 = c(floor_ep1, ceiling_ep1)
range_ep3
```

If a fair die is rolled 1,000 times, the likely amount of occurences for 6 should be between 143 and 190 times. We will reject Ho if 6 is drawn above 190 times. *should we use 95% to standardize with other trials?*

Below is the graphical binomial distribution of EP1.
```{r}
success <- 0:1000

plot(success,dbinom(success,size=1000,prob=.1667),
     type='h',
     main='Binomial Distribution (n = 1000, p = 1/6)',
     ylab='Probability',
     xlab ='# Occurences',
     lwd=3)
```

Hypothesis test EP1: 
```{r}
dice.roll <- sample(1:6, size = 1000, replace = TRUE)
test_ep1 <- table(dice.roll)
test_ep1
```

Given an outcome of 181 occurences of 6, we accept the null hypothesis for EP1. All other sides had an outcome within the range of 143 and 190. However, it was a close call, so we will run the test again, with a 90% confidence to see if 1000 rolls is a large enough number to tighten our range.

Hypothesis test EP1 2.0, with confidence interval 90%.
```{r}
z_90 = 1.645
distance_ep1 = se_ep1 * z_90
floor_ep1 = occurences_ep1 - distance_ep1
ceiling_ep1 = occurences_ep1 + distance_ep1
range_ep1 = c(floor_ep1, ceiling_ep1)
range_ep1
```

Our first test for EP1 would've still been accepted (barely though). We will roll our dice again to see if it stands of another test.
```{r}
dice.roll <- sample(1:6, size = 1000, replace = TRUE)
test_ep1 <- table(dice.roll)
test_ep1
```

Indeed, in this test, we still accept the null hypothesis for number 6. However, 5 rolled a number of occurences below our 90% confidence interval. This shows that this sample number is still too small and overrepresents the theoretical mean.


______


EP2
For EP2, we will repeat the procedure of for EP1 protocols, while changing rolls to 10,000 to test the confidence level of 90% which seemed to be too conservative for 1,000 rolls.

Below we determine expected occurences of 6 in a trial of 10,000 rolls.
```{r}
n_ep2 = 10000
occurences_ep2 = theoretical_prob * n_ep2
occurences_ep2
```
Now, we determine standard error for this trial.
```{r}
se_ep2 = sqrt(n_ep2*(theoretical_prob*(1-theoretical_prob)))
se_ep2
```

In order to determine the range of fair occurences of 6 with confidence level of 90%, we use a z-score of 1.645.
```{r}
z_90 = 1.645
distance_ep2 = se_ep2 * z_90
floor_ep2 = occurences_ep2 - distance_ep2
ceiling_ep2 = occurences_ep2 + distance_ep2
range_ep2 = c(floor_ep2, ceiling_ep2)
range_ep2
```

If a fair die is rolled 10,000 times, the likely amount of occurences for 6 should be between 1,605 and 1728 times. We will reject Ho if 6 is drawn above 1728 times. *should we use 95% to standardize with other trials?*

Below is the graphical binomial distribution of EP1.
```{r}
success <- 0:10000

plot(success,dbinom(success,size=10000,prob=.1667),
     type='h',
     main='Binomial Distribution (n = 10000, p = 1/6)',
     ylab='Probability',
     xlab ='# Occurences',
     lwd=3)
```

Hypothesis test EP2: 
```{r}
dice.roll <- sample(1:6, size = 10000, replace = TRUE)
test_ep2 <- table(dice.roll)
test_ep2
```

Given an outcome of 1716 occurences of 6, we accept the null hypothesis for EP2. All other sides had an outcome within the range of 1605 and 1728. At a 90% confidence, our numbers are still close to the range floor and ceiling. 

______


EP3

For EP3, we are increasing rolls to 100,000 to test the confidence level of 90% and 68%.

Below we determine expected occurences of 6 in a trial of 100,000 rolls.
```{r}
n_ep3 = 100000
occurences_ep3 = theoretical_prob * n_ep3
occurences_ep3
```
Now, we determine standard error for this trial.
```{r}
se_ep3 = sqrt(n_ep3*(theoretical_prob*(1-theoretical_prob)))
se_ep3
```

In order to determine the range of fair occurences of 6 with confidence level of 90%, we use a z-score of 1.645.
```{r}
z_90 = 1.645
distance_ep3 = se_ep3 * z_90
floor_ep3 = occurences_ep3 - distance_ep3
ceiling_ep3 = occurences_ep3 + distance_ep3
range_ep3 = c(floor_ep3, ceiling_ep3)
range_ep3
```

If a fair die is rolled 100,000 times, the likely amount of occurences for 6 should be between 16,473 and 16,861 times. We will reject Ho if 6 is drawn above 16,861 times. *should we use 95% to standardize with other trials?*

Below is the graphical binomial distribution of EP3.
```{r}
success <- 0:100000

plot(success,dbinom(success,size=100000,prob=.1667),
     type='h',
     main='Binomial Distribution (n = 10000, p = 1/6)',
     ylab='Probability',
     xlab ='# Occurences',
     lwd=3)
```

Hypothesis test EP3: 
```{r}
dice.roll <- sample(1:6, size = 100000, replace = TRUE)
test_ep3 <- table(dice.roll)
test_ep3
```

Given an outcome of 16856 occurences of 6, we accept the null hypothesis for EP3. All other sides had an outcome within the range of 16,473 and 16,861. We will repeat this twice to verify.

Hypothesis test EP3 2.0: 
```{r}
dice.roll <- sample(1:6, size = 100000, replace = TRUE)
test_ep3 <- table(dice.roll)
test_ep3
```

Given an outcome of 16552 occurences of 6, we accept the null hypothesis for EP3. All other sides had an outcome within the range of 16,473 and 16,861.

Hypothesis test EP3 3.0: 
```{r}
dice.roll <- sample(1:6, size = 100000, replace = TRUE)
test_ep3 <- table(dice.roll)
test_ep3
```

Given an outcome of 16694 occurences of 6, we accept the null hypothesis for EP3. All other sides had an outcome within the range of 16,473 and 16,861.

Conclusion:
The rule of large numbers hold for our experimental protocols. Increasing the sample size gave us a guarantee that our range of outputs would be closer to the theoretical mean of 16.7% per side. 